### 2017-08-30 revised-chapter.tex
Formatting changes
 - Moved OMNI to subsection of case study

\documentclass{article}

\usepackage{amsmath}
\usepackage{txfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{url}
\usepackage{natbib}
\usepackage{rotating}
\usepackage{adjustbox}
\graphicspath{{figures/}}
\bibliographystyle{biblio.bst}

\title{Probabilistic Forecasting of Geomagnetic Indices using Gaussian Process Models}

\begin{document}
\maketitle

\begin{abstract}

  In this chapter, we give the reader an in depth view into building of
  probabilistic forecasting models for geomagnetic time series using the
  \emph{Gaussian Process} methodology outlined in the previous
  chapters. We highlight design decisions and practical issues that must
  be addressed in order to use \emph{Gaussian Process} models for
  probabilistic prediction of a quantity of interest. As a pedagogical
  example, we formulate, train and test a family of \emph{Gaussian
    Process} Auto-Regressive models for one hour ahead prediction of
  the $Dst$ geomagnetic index.

\end{abstract}

\section{Geomagnetic Time Series \& Forecasting}

The Earth's magnetosphere is the region above the ionosphere that is
defined by the extent of the Earth's magnetic field in space. It
extends several tens of thousands of kilometers into space, a region
which is impinged upon constantly by the solar wind. Ionized plasma
ejected by the sun couples with the Earth's magnetic field in a
complex manner leading to highly non-linear and chaotic processes
which determine the state of the magnetosphere. It is quite common to
reduce these complex dependencies by condensing the state of the
Earth's magnetosphere into a set of geomagnetic indices.

Geomagnetic indices come in various forms, they may take continuous or
discrete values and may be defined with varying time
resolutions. Their values are often calculated by averaging or
combining a number of readings taken by instruments, usually
magnetometers, around the Earth. Each geomagnetic index is a proxy for
a particular kind of phenomenon. Some common indices are the $Kp$,
$Dst$ and the $AE$ index.

\begin{enumerate}
\item $Kp$: The Kp-index is a discrete valued global geomagnetic
  activity index and is based on 3 hour measurements of the
  K-indices \citep{Bartels}. The K-index itself is a three hour
  long quasi-logarithmic local index of the geomagnetic activity,
  relative to a quiet day curve for the given location.

\item $AE$: The Auroral Electrojet Index, $AE$, is designed to
  provide a global, quantitative measure of auroral zone magnetic
  activity produced by enhanced ionospheric currents flowing below
  and within the auroral oval \citep{AEIndex}. It is a continuous
  index which is calculated every hour.

\item $Dst$: A continuous hourly index which measures the
  magnitude of the magnetic field induced by near equatorial ring
  currents \citep{DesslerAndParker}. A large negative value of
  $Dst$ ($ \leq -100 \ nT$) indicates occurrence of geomagnetic
  storm events which are of particular importance in space weather
  prediction due to their adverse effects on telecommunications
  infrastructure, navigation systems, power grids, satellites etc. 
\end{enumerate}

Space weather forecasting systems usually use in-situ measurements of
solar wind parameters, taken by satellites, as well as historical data
of indices to produce forecasts for various geomagnetic time
series. Real time data of solar wind parameters are recorded by the \emph{ACE}
and \emph{DISCOVR} missions. Geomagnetic indices are often recorded in
ground based measurements. The the \emph{Space Physics Data Facility}
(SPDF) hosted by NASA is a gateway for accessing data from a variety
of sources, including those mentioned above.

In this chapter for the purpose of exposition, we will focus on one
hour ahead prediction of the $Dst$ time series. From the data we will
use time histories of $Dst$, solar wind speed $V$ and the
$z$-component of the interplanetary magnetic field $B_z$. Our choice
is motivated by both physical and practical considerations: solar wind
speed and interplanetary magnetic field are important external drivers
for particle injection into the magnetosphere and they have large
coverage in the \emph{OMNI} data with relatively lesser number of
missing records.

\section{$Dst$ Forecasting}

\subsection{Models \& Algorithms}

A number of modeling techniques have been applied for the prediction
of the $Dst$ index. One of the earliest forecasting techniques
involves calculating the $Dst(t)$ as a solution of an \emph{Ordinary
  Differential Equation} (ODE) which expressed the rate of change of
$Dst(t)$ as a combination of two terms: decay and injection $\frac{d
  Dst(t)}{dt} = Q(t) - \frac{Dst(t)}{\tau}$, where $Q(t)$ relates to
the particle injection from the plasma sheet into the inner
magnetosphere. This method was presented first by \citet{JGR:JGR10260}
and later modified and extended in works such as \citet{Wang:Dst},
\citet{JGRA:JGRA14856}, \citet{Ballatore2014} and others.

% Talk about NARMAX Dst
Important empirical geomagnetic prediction models include the
\emph{Nonlinear Auto-Regessive Moving Average with eXogenous inputs}
(NARMAX) methodology (see \citet{doi:10.1080/00207178908559767},
\citet{GRL:GRL13494}, \citet{GRL:GRL20944}, \citet{JGRA:JGRA18657},
\citet{balikhin:narmax}, \citet{JGRA:JGRA20661},
\citet{JGRA:JGRA50192}) and \emph{Artificial Neural Networks} (ANN)
based models (\citet{Lund}, \citet{JGRA:JGRA17461},
\citet{SWE:SWE286}, \citet{pallocchia:hal-00318011}) for time series
prediction of $Dst$ and $Kp$ indices from interplanetary magnetic
field data and solar wind parameters.

\subsection{Probabilistic Forecasting}
% Talk about need for probabilistic forecasts.ny
Although much research has been done on prediction of the $Dst$ index,
much less has been done on probabilistic forecasting of $Dst$. One
such work described in \citet{McPherron:2013} involves identification
of high speed solar wind streams using the WSA model, using
predictions of high speed streams to construct ensembles of $Dst$
trajectories which yield the quartiles of $Dst$ time series.

Probabilistic forecasting is of particular importance in geophysics
applications as the end users of forecasts often require confidence
bounds on the said forecasts. It is in this context where
\emph{Gaussian Processes} become especially attractive due to their
inherent probabilistic formulation and tractability of exact inference.

\section{Gaussian Processes}

\emph{Gaussian Processes} first appeared in machine learning research
in \citet{Neal:1996:BLN:525544}, as the limiting case of Bayesian
inference performed on neural networks with infinitely many neurons in
the hidden layers. Although their inception in the machine learning
community is recent, their origins can be traced back to the
geo-statistics research community where they are known as
\emph{Kriging} methods \citep{krige1951statistical}. In pure
Mathematics, \emph{Gaussian Processes} have been studied extensively
and their existence was first proven by Kolmogorov's extension theorem
\citep{tao2011introduction}. The reader is referred to
\citet{Rasmussen:2005:GPM:1162254} for an in depth treatment of
Gaussian Processes in machine learning. In space weather forecasting
literature \emph{Gaussian Process Regression} has been used for one
hour ahead prediction of the Dst index \citep{SWE:SWE20479}.

Without going into too many details, we give a quick recap of the
formulation and exact inference in \emph{Gaussian Process Regression} (GPR) models.

\subsection{Gaussian Process Regression: Formulation}

Our aim is to infer an unknown function $f(\mathbf{x})$ from its noise
corrupted measurements $(\mathbf{x}_i, y_i)$ where $y_i =
f(\mathbf{x}_i) + \epsilon$ and $\epsilon \sim \mathcal{N}(0,
\sigma^2)$ is independent and identically distributed Gaussian noise.

A \emph{Gaussian Process} model represents the finite dimensional
probability distribution of $f(\mathbf{x}_i)$ by a multivariate
gaussian having a particular structure for its mean and covariance as
shown in \ref{eq:normal} - \ref{eq:sto}.

\begin{align}
  \mathbf{f} = & \left( \begin{array}{c} f(\mathbf{x}_1) \\ f(\mathbf{x}_2) \\ \vdots \\ f(\mathbf{x}_N) \end{array} \right) \label{eq:fvalues}\\
  \vspace{2\baselineskip}
  \mathbf{f} | \mathbf{x}_1, \cdots, \mathbf{x}_N \sim & \mathcal{N}\left( \mathbf{m}, \mathbf{K} \right)  \label{eq:normal}\\
  \vspace{2\baselineskip}
  \mathbb{P}( \mathbf{f} \ | \ \mathbf{x}_1, \cdots, \mathbf{x}_N) = & \frac{1}{(2\pi)^{n/2} det(\mathbf{K})^{1/2}} exp \left(-\frac{1}{2} (\mathbf{f} - \mathbf{m})^T \mathbf{K}^{-1} (\mathbf{f} - \mathbf{m}) \right) \label{eq:sto}
\end{align}

In order to uniquely define the distribution of $\mathbf{f}$, it is
required to specify the mean $\mathbf{m}$ and the covariance matrix
$\mathbf{K}$. For this probability density to be valid, there are
further requirements imposed on $\mathbf{K}$:

\begin{enumerate}
\item Symmetry: $\mathbf{K}_{ij} = \mathbf{K}_{ji} \ \forall i,j \in {1, \cdots, N} $ 
\item Positive Semi-definiteness: $\mathbf{z}^T \mathbf{K} \mathbf{z} \geq 0 \ \forall \mathbf{z} \in \mathbb{R}^N$  
\end{enumerate}

In \emph{Gaussian Processes} the individual elements of $\mathbf{x}$
and $\mathbf{K}$ are specified in the form of functions as shown
below.

\begin{align}
  \mu_i = & \mathbb{E}[f(\mathbf{x}_i)] := m(\mathbf{x}_i) \\
  \Lambda_{ij} = & \mathbb{E}[(f(\mathbf{x}_i) - \mu_i)(f(\mathbf{x}_j) - \mu_j)] := K(\mathbf{x}_i, \mathbf{x}_j)
\end{align}

In the machine learning community, $m(.)$ and $K(.,.)$ are known as
the \emph{mean function} and \emph{covariance function} or
\emph{kernel function} of the process respectively. Giving a closed
form expression for $m(.)$ and $K(.,.)$ uniquely specifies a
particular \emph{Gaussian Process} and so a GP model is often
expressed in the following notation.

\begin{equation}
  f(\mathbf{x}) \sim \mathcal{GP}(m(\mathbf{x}), K(\mathbf{x}, \mathbf{x}'))
\end{equation}

\subsection{Gaussian Process Regression: Inference}

In order to generate predictions $f(\mathbf{x}^{*}_i)$ for a set of
test points $ {\mathbf{x}^{*}_i : \forall i \in 1, \cdots, M} $. Using
the multivariate Gaussian distribution in equation (\ref{eq:sto}) we
can construct the joint distribution of $f(\mathbf{x})$ over the
training and test points.

\begin{align}
  \mathbf{f}_* = & \left( \begin{array}{c} f(\mathbf{x^{*}_1}) \\ f(\mathbf{x^{*}_2}) \\ \vdots \\ f(\mathbf{x^{*}_M}) \end{array} \right)_{M \times 1} \\
  \vspace{4\baselineskip}
  \left( \begin{array}{c} \mathbf{y} \\ \mathbf{f_*} \end{array} \right) | \ \ \mathbf{X}, \mathbf{X}_* \sim & 
                                                                                                               \mathcal{N}\left(\mathbf{0}, \left[ \begin{array}{cc} \mathbf{K} + \sigma^{2} \mathbf{I} & \mathbf{K}_{*} \\ \mathbf{K}_{*}^T & \mathbf{K}_{**} \end{array} \right ] \right) \label{eq:dist}
\end{align}

Probabilistic predictions $f_*$ can be generated by constructing the
conditional distribution
$\mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*}$ which is also a
multivariate gaussian as shown in equation \ref{eq:posterior}.

\begin{align}
  & \mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*} \sim \mathcal{N}(\mathbf{\bar{f}_*}, \Sigma_*)  \label{eq:posterior} \\
  & \mathbf{\bar{f}_*} =  \ \mathbf{K}^T_{*} [\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1} \mathbf{y} \label{eq:posteriormean} \\
  & \Sigma_* = \ \mathbf{K}_{**} - \mathbf{K}^T_{*} \left(\mathbf{K} + \sigma^{2} \mathbf{I}\right)^{-1} \mathbf{K}_{*} \label{eq:posteriorcov}
\end{align}

\section{One Hour Ahead $Dst$ Prediction} \label{sec:osa}

\subsection{Data Source: OMNI}
The \emph{OMNI} data set is a collation of in-situ measurements taken
by orbiting satellites and ground based measurements of geomagnetic
indices. The data itself is represented as time averaged versions of
the various quantities and is available at two frequencies, by hour
and by minute. The quantities recorded can be broadly classified into
two groups.

\begin{enumerate}

\item Plasma Parameters and Magnetic fields: Measurements of the solar
  wind particularly the solar wind speed, flow angles, pressure,
  proton temperature, proton density and field measurements such as
  strength of electric and magnetic fields as well as their components
  in various geomagnetic coordinate systems.

\item Indices: $Kp$, $Ae$, $Ap$, $Dst$, sunspot number.

\end{enumerate}

To get a more indepth understanding about the \emph{OMNI} data,
interested readers are encouraged to visit the NASA omni web service (\url{https://omniweb.gsfc.nasa.gov}).

\subsection{Gaussian Process $Dst$ Model}

Below in equations (\ref{eq:Dst}) - (\ref{eq:DstGP}) we outline a
\emph{Gaussian Process} formulation for \emph{OSA} prediction of
$Dst$. A vector of features $\mathbf{x}_{t-1}$ is used as input to an
unknown function $f(\mathbf{x}_{t-1})$.

The features $\mathbf{x}_{t-1}$ can be any collection of quantities in
the hourly resolution OMNI data set. Generally $\mathbf{x}_{t-1}$ are
time histories of $Dst$ and other important variables such as plasma
pressure $p(t)$, solar wind speed $V(t)$, $z$ component of the
interplanetary magnetic field $B_z(t)$.

\begin{align}
  Dst(t) & =  f(\mathbf{x}_{t-1}) + \epsilon \label{eq:Dst} \\
  \epsilon & \sim  \mathcal{N}(0, \sigma^2) \label{eq:GPNoise} \\
  f(x_t) & \sim  \mathcal{GP}(m(\mathbf{x}_t), K_{osa}(\mathbf{x}_t, \mathbf{x}_s)) \label{eq:DstGP} \\
\end{align}

We consider two choices for the input features $\mathbf{x}_{t-1}$
leading to two variants of \emph{Gaussian Process} regression for
$Dst$ time series prediction.

\subsection{Gaussian Process Auto-Regressive (GP-AR)} \label{sec:gpar}

The simplest auto-regressive models for \emph{OSA} prediction of $Dst$
are those that use only the history of $Dst$ to construct input
features for model training. The input features $\mathbf{x}_{t-1}$ at
each time step are the history of $Dst(t)$ until a time lag of $p$ hours.

\begin{align*}
  \mathbf{x}_{t-1} = & \left(Dst(t-1), \cdots , Dst(t-p+1)\right)
\end{align*}

\subsection{Gaussian Process Auto-Regressive with eXogenous inputs (GP-ARX)} \label{sec:gparx}

Auto-regressive models can be augmented by including exogenous
quantities in the inputs $\mathbf{x}_{t-1}$ at each time step, in
order to improve predictive accuracy. $Dst$ gives a measure of ring
currents, which are modulated by plasma sheet particle injections into
the inner magnetosphere during substorms. Studies have shown that the
substorm occurrence rate increases with solar wind velocity (high
speed streams) \citet{Kissinger2011,Newell2016}. Prolonged southward
interplanetary magnetic field (IMF) $z$-component ($B_z$) is needed
for sub-storms to occur \citet{McPherron1986}. An increase in the
solar wind electric field, $V_{sw}B_z$, can increase the dawn-dusk
electric field in the magnetotail, which in turn determines the amount
of plasma sheet particle that move to the inner magnetosphere
\citet{Friedel2001}. Therefore, our exogenous parameters consist of
solar wind velocity $V_{sw}$ and IMF $B_z$.

In this model we choose distinct time lags $p$, $p_{v}$ and $p_{b}$
for $Dst$, $V$ and $B_z$ respectively.

\begin{align*}
  \mathbf{x}_{t-1} & = (Dst(t-1), \cdots , Dst(t-p+1), \\
                   & \ \ \ \ \  V_{sw}(t-1), \cdots, V_{sw}(t-p_{v}+1),\\
                   & \ \ \ \ \  B_{z}(t-1), \cdots, B_{z}(t-p_{b}+1))
\end{align*}

\section{One Hour Ahead $Dst$ Prediction: Model Design}\label{sec:modeldesign}

\subsection{Choice of Mean Function}

Mean functions in GPR models encode trends in the data, they are the
baseline predictions the model falls back to in case the training and
test data have little correlation as predicted by the kernel
function. If there is no prior knowledge about the function to be
approximated, \citet{Rasmussen:2005:GPM:1162254} state that it is
perfectly reasonable to choose $m(\mathbf{x} = 0)$ as the mean
function, as long as the target values are normalized. In the case of
the $Dst$ time series, it is known that the so called
\emph{persistence model} $\hat{D}st(t) = Dst(t-1)$ performs quite well
in the context of OSA prediction. We therefore choose the
\emph{persistence model} as the mean function in our OSA Dst models.

\subsection{Choice of Kernel}

For the success of a \emph{Gaussian Process} model an appropriate
choice of kernel function is paramount. The symmetry and positive
semi-definiteness of \emph{Gaussian Process} kernels implies that they
represent inner-products between some basis function representation of
the data. The interested reader is suggested to refer to
\citet{Berlinet2004}, \citet{Scholkopf:2001:LKS:559923} and
\citet{hofmann2008} for a thorough treatment of kernel functions and
the rich theory behind them. Some common kernel functions used in
machine learning are listed in table \ref{table:kernel}.

\begin{table}[h]
  \caption{Popular Kernel functions used in GPR models}
  \centering
  \begin{tabular}{l c c}
    \hline
    Name  & Expression & Hyperparameters  \\
    \hline
    Radial Basis Function (RBF)  & $\frac{1}{2} exp(-||\mathbf{x} - \mathbf{y}||^2/l^2)$  & $l \in \mathbb{R}$   \\

    Polynomial  & $(\mathbf{x}^\intercal \mathbf{y} + b)^d$ & $b \in \mathbb{R}, d \in \mathbb{N}$   \\

    Laplacian  & $exp(-||\mathbf{x} - \mathbf{y}||_{1}/\theta)$  & $\theta \in \mathbb{R}^+$  \\

    Student's T  & $1/(1 + ||\mathbf{x} - \mathbf{y}||_{2}^d)$ & $d \in \mathbb{R}^{+}$\\

    Maximum Likelihood Perceptron  & $sin^{-1}(\frac{w\mathbf{x}^\intercal \mathbf{y} + b}{\sqrt{w\mathbf{x}^\intercal \mathbf{x} + b + 1} \sqrt{w\mathbf{y}^\intercal \mathbf{y} + b + 1}})$ & $w, b \in \mathbb{R}^{+}$\\
    \hline
  \end{tabular}
  \label{table:kernel}
\end{table}

The quantities $l$ in the \emph{Radial Basis Function kernel}, and $b$
and $d$ in the polynomial kernel are known as
\emph{hyperparameters}. Hyperparameters give flexibility to a
particular kernel structure, for example $d = 1, 2, 3, \cdots$ in the
polynomial kernel represents linear, quadratic, cubic and higher order
polynomials respectively. The process of assigning values to the
\emph{hyperparameters} is crucial in the model building process and is
known as \emph{model selection}.

In this text, we construct GPR models with a combination of the
\emph{maximum likelihood perceptron} kernel and \emph{student's T}
kernel as shown in equations (\ref{eq:usedKernel}). The \emph{maximum
  likelihood perceptron} kernel is the \emph{Gaussian Process}
equivalent of a single hidden layer feed-forward neural network model
as demonstrated in \citet{Neal:1996:BLN:525544}.

\begin{align}
  K_{osa}(\mathbf{x}, \mathbf{y}) & = K_{mlp}(\mathbf{x}, \mathbf{y}) + K_{st}(\mathbf{x}, \mathbf{y}) \label{eq:usedKernel} \\
  K_{mlp}(\mathbf{x}, \mathbf{y}) & = sin^{-1}(\frac{w\mathbf{x}^\intercal \mathbf{y} + b}{\sqrt{w\mathbf{x}^\intercal \mathbf{x} + b + 1} \sqrt{w\mathbf{y}^\intercal \mathbf{y} + b + 1}}) \\
  K_{st}(\mathbf{x}, \mathbf{y}) & = \frac{1}{1 + ||\mathbf{x} - \mathbf{y}||_{2}^d}
\end{align}

\subsection{Model Selection: Hyperparameters}

Given a GPR model with a kernel function $K_\theta$, the problem of
model selection consists of finding appropriate values for the kernel
hyperparameters $\theta = \left(\theta_1, \theta_2, \cdots,
  \theta_i\right)$. In order to assign a value to $\theta$, we must
define an objective function ($Q(\theta)$) which represents our
confidence that the GP model built from a particular value of $\theta$
is the best performing model. Since GP models are constructed from
assumptions about the conditional probability distribution of the data
$p(\mathbf{y}|\mathbf{X})$, it is natural to use the negative
log-likelihood of the training data as a model selection criterion.

\begin{align*}
  Q(\theta) & = - log \ p(\mathbf{y}|\mathbf{X}, \theta) \\
            & = \frac{1}{2} \mathbf{y}^\intercal (\mathbf{K}_\theta + \sigma^{2} \mathbf{I})^{-1} \mathbf{y} + \frac{1}{2}|\mathbf{K}_\theta + \sigma^{2} \mathbf{I}| + \frac{N}{2}log(2\pi) \\
  \mathbf{K}_\theta & = [K_{\theta}(\mathbf{x}_i, \mathbf{x}_j)]_{N \times N}
\end{align*}

Since a GP model expresses prior assumptions about the finite
dimensional distribution of the quantitiy of interest, the expression
for $Q(\theta)$ follows directly from the expression of the
probability density function of a multivariate gaussian.

The model selection problem can now be expressed as the minimization
problem shown below.

\begin{align*}
  \theta^* = arg\min_{\theta} \ Q(\theta)
\end{align*}

\citet{Rasmussen:2005:GPM:1162254} note that there are ready
interpretations to the three terms which add up to give $Q(.)$. The
first term $\frac{1}{2} \mathbf{y}^\intercal (\mathbf{K}_\theta +
\sigma^{2} \mathbf{I})^{-1} \mathbf{y}$ quantifies the fit of the
model with respect to the training data, the second term
$\frac{1}{2}|\mathbf{K}_\theta + \sigma^{2} \mathbf{I}|$ is a
complexity penalty, in non-parametric models such as \emph{Gaussian
  Processes} the model complexity grows with the number of training
data points $N$. The third term $\frac{N}{2}log(2\pi)$ is a
normalization term.

The objective function $Q(\theta)$ can have multiple local minima, and
evaluating the value of $Q(.)$ at any given $\theta$ requires
inversion of the matrix $\mathbf{K}_\theta + \sigma^{2} \mathbf{I}$
which has a time complexity $O(N^3)$ as noted above. In the interest
of saving computational cost, one cannot use exhaustive search through
the domain of the hyperparameters to inform our choice for $\theta$.

The model selection problem has historically received less attention
due to a combination of difficulty and numerous design decisions like
choice of hyper-parameter optimization techniques, which makes model
selection as much of an art as science.

Choosing a suitable heuristic or algorithm for model selection for GP
models is quite often a balancing exercise between the computational
limitations of the modeler and the need to explore the hyper-parameter
space in a manner that yields a satisfactory predictive model. Some of
the techniques used for model selection in the context of GPR include
the following.

\subsubsection*{Grid Search}

This is the simplest procedure that may be applied for performing
model selection in GPR models. This routine constructs a grid of
values for $\theta$ as the cartesian product of one dimensional grids
for each $\theta_i$, evaluate $Q(.)$ at each such grid point and
choose the configuration which yields minimum value of $Q(.)$. The
advantage of this technique is its simplicity and control over the
computational cost of training GP models for practical problems. The
computational cost of this procedure scales with the total number of
points on the grid, this is controlled by the scales and number of
grid points for each hyper-parameter $\theta_i$.

\subsubsection*{Coupled Simulated Annealing}

Introduced in \citet{Xavier-De-Souza2010}, \emph{Coupled Simulated
  Annealing} (CSA) is a family of optimization techniques which is a
generalization of \emph{Simulated Annealing} (SA)
\citep{Kirkpatrick671}. The CSA optimization technique can be
understood as a set of parallel SA processes which are coupled by
their acceptance probabilities.

From an implementation point of view, CSA follows the same procedure
as \emph{grid search}, but after evaluation of the objective $Q(.)$ on
the grid, each grid point is iteratively mutated in a random walk
fashion. This mutation is accepted or rejected according to the new
value of $Q(.)$ as well as its value on the other grid points. This
procedure is iterated until some stop criterion is reached.

To illustrate how CSA updates are made and accepted consider a
population or ensemble of hyper-parameter values ${x_i}$ which can be
considered to be initialised in a grid based fashion. For each
configuration $x_i$ a probing solution $y_i = x_i + \epsilon$ is
generated where $\epsilon$ is drawn from some random variable. The
acceptance probability of the solution is calculated as shown below.

\begin{align*}
  \mathbb{P}(x_i \rightarrow y_i | {x_i}) & = \frac{exp(-E(y_i)/T)}{exp(-E(y_i)/T) + \gamma} \\
  \gamma & = \sum_{j}{exp \left ( \frac{-E(x_j)}{T} \right )}
\end{align*}

The quantity $\gamma$ defines the \emph{coupling} between the states
${x_i}$ of the ensemble, when the number of configurations in the
ensemble $x_i$ is one this rule reduces to the classical
\emph{Metropolis-Hastings} acceptance rule of the SA algorithm. Up to
4 variants of the CSA algorithm have been defined with different
acceptance rules and temperature annealing schedules, the reader is
advised to refer to \citet{Xavier-De-Souza2010} for an in-depth
treatment of the CSA optimization class.

\subsubsection*{Maximum Likelihood}

This technique, as described in \citet{Rasmussen:2005:GPM:1162254}, is
a form of \emph{gradient descent}. It involves starting with an
initial guess for $\theta$ and iteratively improving it by calculating
the gradient of $Q(.)$ with respect to $\theta$. The gradient of
$Q(.)$ with respect to each hyper-parameter $\theta_i$ can be
calculated analytically as shown below.

\begin{align*}
  \frac{\partial}{\partial \theta_i} log \ p(\mathbf{y}|\mathbf{X}, \theta) & = \frac{1}{2} \mathbf{y}^\intercal K^{-1} \frac{\partial K}{\partial \theta_i} K^{-1} \mathbf{y}  - \frac{1}{2} tr(K^{-1} \frac{\partial K}{\partial \theta_i})\\
                                                                            & = \frac{1}{2} tr((\mathbf{\alpha}\mathbf{\alpha}^\intercal - K^{-1}) \frac{\partial K}{\partial \theta_i}) \ \ where \ \mathbf{\alpha} = K^{-1} \mathbf{y}
\end{align*}

Although gradient based \emph{Maximum Likelihood} looks very similar
to the \emph{Gradient Descent} used to learn parametric regression
models like \emph{Feed forward Neural Networks}, in the GPR model
selection context it introduces an extra computational cost of
calculating the gradient of $Q(\theta)$ with respect to each
$\theta_i$ in every iteration. Applying this method can sometimes lead
to overfitting of the GPR model to the training data \citep{Rasmussen:2005:GPM:1162254}.

\subsection{Model Selection: Auto-regressive Order}

Apart from having a practical method for choosing values of kernel
hyperparameters, we must also address the issue of choosing the time
delays $p$ and $p_v, p_b$ for the GP-AR and GP-ARX models. One might
argue that we group these quantities with the kernel hyperparameters
in the model selection procedure, in the context of the GP-AR and
GP-ARX models there is one reason why this might not be beneficial.

The value of kernel hyperparameters $\theta^*$ which maximise the
marginal likelihood $Q(.)$ are conditional on the chosen model orders
$p$ and $p, p_v, p_b$ for GP-AR and GP-ARX respectively. Since the
model orders determine the size of the input space of the predictive
models, they influence the values of characteristic length scales and
other kernel hyperparameters which yield a good fit.

Not only do the model orders influence the value of kernel
hyperparameters $\theta^*$ which maximise model fit, but increasing
values of $p, p_v, p_b$ also make it more difficult for the model
selection algorithms to search the hyper-parameter space, due to the
curse of dimensionality.

An empirical way to choose the values of $p, p_v, p_b$ for the GP-AR
and GP-ARX models is by choosing an independent validation data set
which is used to evaluate the GP models obtained from the model
selection procedures of the previous section. It is then possible for
the modeler to choose these model orders in way to balance performance
and dimensionality of input features.

\section{GP-AR and GP-ARX: Workflow Summary}\label{sec:workflow}

Section \ref{sec:modeldesign} brings together all the practical issues
with respect to building of GP-AR and GP-ARX models for predicting
$Dst$, they can be condensed to form the following steps.

\begin{enumerate}

\item \textbf{Data Preprocessing}: Three seperate data sets for model
  training and hyper-parameter selection, model order selection and
  model evaluation must be created. Care must be taken that the size
  of the training sets does not delay the processes of kernel
  hyper-parameter selection. The training, validation and test data
  sets created from this procedure must all contain non-overlapping
  geomagnetic storm events.

  We selected OMNI data sections 00:00 January 3 2010 to 23:00 January
  23 2010 and 20:00 August 5 2011 to 22:00 August 6 2011 for training
  the GP-AR and GP-ARX models. The first training data section
  consists of ambient fluctuations of $Dst$ while the second contains
  a geomagnetic storm. A set of 24 storm events listed in table
  \ref{table:validationstorms} is kept aside as a validation data set
  for selecting model orders $p, p_v, p_b$ while a list of 63 storm
  events listed in table \ref{table:teststorms} is used for final
  model evaluation.

\item \textbf{Choice of model components}:  Kernel, mean function, model selection procedure.

  We choose a kernel $K_{osa}$ of the form given in equation
  \ref{eq:usedKernel}. In order to find appropriate values of the
  hyperparameters of $K_{osa}$, we apply \emph{grid search},
  \emph{coupled simulated annealing} and \emph{maximum likelihood}
  methods. We fix the parameters $d$ and $\sigma^2$ of $K_{st}$ and
  model noise to values $0.01$ and $0.2$ respectively, the remaining
  parameters $w$ and $b$ are kept free to be calculated by model
  selection. Table \ref{table:modelselection} summarizes the settings
  used to run each model selection procedure.

\item \textbf{Choice of Performance Metrics}: To compare the
  performance of GP and GP-ARX models of varying auto-regressive
  orders, their predictive performance during the validation (and
  later final evaluation) phase must be calculated using some clearly
  defined metrics. We use the following as performance metrics for our
  modeling experiments.

  \begin{enumerate}
  \item The mean absolute error.
    \begin{equation}
      MAE = \sum_{t=1}^{n} \left |(Dst(t) - \hat{D}st(t)) \right | / n
    \end{equation}
  \item The root mean square error.
    \begin{equation}
      RMSE = \sqrt{\sum_{t=1}^{n} (Dst(t) - \hat{D}st(t))^2 / n}
    \end{equation}
  \item Correlation coefficient between the predicted and actual value of $Dst$.
    \begin{equation}
      CC = Cov(Dst, \hat{D}st)/\sqrt{Var(Dst) Var(\hat{D}st)}
    \end{equation}
  \end{enumerate}

\item \textbf{Model Order Selection}: In order to find appropriate
  values for the auto-regressive time lags $p, p_v, p_b$ for $Dst$,
  solar wind speed and IMF $B_z$ respectively, create models upto a
  certain maximum time lag i.e. $3 \leq p \leq 12$ for GP-AR and $3
  \leq p + p_v + p_b \leq 12$ for GP-ARX. For each such model perform
  the model selection routine to select values of the kernel
  hyperparameters and evaluate the resultant model on the validation
  set. Choose the value of $p, p_v, p_b$ which achieves the predictive
  performance on the validation set (i.e. minimizes the \emph{mean
    absolute error}).

\item \textbf{Final Evaluation}: After selecting GP-AR and GP-ARX
  models with particular values of auto-regressive orders and kernel
  hyperparameters in the model selection phase, the best performing
  GP-AR and GP-ARX models are then evaluated on the test data set in
  order to get an unbiased picture of their predictive performance. In
  our case we use the already chosen test set outlined in table
  \ref{table:teststorms} for final model evaluation. The GP based OSA
  prediction models are compared against the naive \emph{Persistence}
  model which serves as a performance baseline for the OSA prediction
  of $Dst$.

\end{enumerate}

\section{Practical Issues: Software}

Scientific computing, whether it be in the form of simulation, data
analysis or modeling is brought to fruition through well thought
choices and usage of software packages. \emph{Gaussian Processes} have
been implemented in a number software distributions, we discuss two of
them here.

\begin{enumerate}

\item \textbf{PlasmaML}: An open source project
  \url{https://gitlab.com/MD-CWI-NL/PlasmaML} maintained at the CWI
  Amsterdam, it is a software toolbox for Machine Learning
  applications in Space \& Plasma Physics. It is comprised of a number
  of modules, we use codes from the \textit{omni} module for the
  experiments and results shown in section \ref{sec:results}. It
  leverages the DynaML machine learning toolkit
  \url{https://github.com/transcendent-ai-labs/DynaML} which has a
  well developed library for building GP and Kernel based models and
  also includes in built support for a number of kernel functions as
  well as the \emph{Grid Search}, \emph{Maximum Likelihood} and
  \emph{Coupled Simulated Annealing} model selection routines applied
  in this chapter. \textit{PlasmaML} is written in the \textit{Scala}
  programming language making it distributable as a \emph{Java Runtime
    Environment} \textit{jar} file.

\item \textbf{GPML} \citep{GPML}: A \emph{MATLAB} based toolbox for
  building GP models which has support for a variety of features. The
  software includes a number of kernel functions supported off the
  shelf, approximate as well as exact inference techniques and support
  for non-Gaussian likelihood functions. 
\end{enumerate}

\section{Experiments \& Results}\label{sec:results}

After running the workflow steps as outlined in section
\ref{sec:workflow} above, we are at a position to analyse the
performance of the GP-AR and GP-ARX family of models on task of OSA
$Dst$ prediction.

\subsection*{Model Selection \& Validation Performance}

The model validation workflow described in section \ref{sec:workflow}
produces data containing model validation scores for GP-AR and GP-ARX
models with different values of autoregressive orders $p_t$. In order
to better understand the over all trend, we group the performance
scrores by unique values of $p_t = p + p_v + p_b$ and analyse the
summary statistics with increasing $p_t$

Figures \ref{fig:CompareMae} and \ref{fig:CompareCC} show how the mean
absolute error and coefficient of correlation as calculated on the
validation set storm events of table \ref{table:validationstorms},
vary with increasing model order for GP-AR and GP-ARX. The results are
represented as box and whisker plots, in which a rectangle is drawn to
represent the first and third quartiles, with a horizontal line inside
to indicate the median value, outlying points are shown as dots while
the whiskers indicate the smallest and largest non-outliers. In both
cases, the predictive performance first improves and then stagnates or
worsens with increasing model order.

\begin{figure}[h]
  \includegraphics[width=\textwidth]{Compare-mae.png}
  \caption{Mean Absolute Error on validation set storms vs model order for GP-AR and GP-ARX. \textbf{Key}: Rectangle borders represent the first and third quartiles, with a horizontal line inside to indicate the median value, outlying points are shown as dots and whiskers indicate the smallest and largest non-outliers}
  \label{fig:CompareMae}
\end{figure}

\begin{figure}[h]
  \includegraphics[width=\textwidth]{Compare-cc.png}
  \caption{Coefficient of Correlation on validation set storms vs model order for GP-AR and GP-ARX. \textbf{Key}: Rectangle borders represent the first and third quartiles, with a horizontal line inside to indicate the median value, outlying points are shown as dots and whiskers indicate the smallest and largest non-outliers}
  \label{fig:CompareCC}
\end{figure}

From the charts above, we can choose the model order which yields the
best \emph{mean absolute error}, for GP-AR it is $p_t = 6$ while for
GP-ARX it is $p_t = 11$. Further examination of the validation results
shows that in the scheme $p_t = 11$ choosing $p = 7, p_v = 1, p_b = 3$ gives superior results.

\subsection*{Comparison of Hyper-parameter Selection Algorithms}
Figures \ref{fig:CompareMaeARX} and \ref{fig:CompareCCARX} break down
the results for GP-ARX by the model selection routine used. Apart from
the general trend observed in \ref{fig:CompareMae} and
\ref{fig:CompareCC}, we also observe that \emph{grid search} and
\emph{coupled simulated annealing} give superior performance as
compared to gradient based \emph{maximum likelihood}.

\begin{figure}[h]
  \includegraphics[width=\textwidth]{Compare-mae-arx.png}
  \caption{Mean Absolute Error on validation set storms vs model order for GP-AR and GP-ARX for \emph{CSA}, \emph{GS} and \emph{ML} model selection routines. \textbf{Key}: Rectangle borders represent the first and third quartiles, with a horizontal line inside to indicate the median value, outlying points are shown as dots and whiskers indicate the smallest and largest non-outliers}
  \label{fig:CompareMaeARX}
\end{figure}

\begin{figure}[h]
  \includegraphics[width=\textwidth]{Compare-cc-arx.png}
  \caption{Coefficient of Correlation on validation set storms vs model order for GP-AR and GP-ARX for \emph{CSA}, \emph{GS} and \emph{ML} model selection routines. \textbf{Key}: Rectangle borders represent the first and third quartiles, with a horizontal line inside to indicate the median value, outlying points are shown as dots and whiskers indicate the smallest and largest non-outliers}
  \label{fig:CompareCCARX}
\end{figure}

\subsection*{Final Evaluation}

After choosing the best performing GP-AR and GP-ARX models, we
calculate their performance on the test set of table
\ref{table:teststorms}. The results of these model evaluations are
summarized in table \ref{table:results}, the GP-AR and GP-ARX models
improve upon the performance of the \emph{persistence model}.

\subsection*{Sample Predictions with Error Bars}

Figures \ref{fig:ComparePred1}, \ref{fig:ComparePred2} and
\ref{fig:ComparePred3} show OSA predictions of the GP-ARX model with
$\pm \sigma$ error bars for three storm events in the time period
between 1998 and 2003. The GP-ARX model gives accurate predictions
along with plausible error bars around its mean predictions.

\begin{figure}[h]
  \includegraphics[width=\textwidth]{PredictionsModel1/PredErrBars_Storm43.png}
  \caption{OSA Predictions with $\pm \sigma$ error bars for event: 2003/06/17 to 2003/06/19}
  \label{fig:ComparePred1}
\end{figure}

\begin{figure}[h]
  \includegraphics[width=\textwidth]{PredictionsModel1/PredErrBars_Storm16.png}
  \caption{OSA Predictions with $\pm \sigma$ error bars for event: 2012/03/08 to 2012/03/10}
  \label{fig:ComparePred2}
\end{figure}

\begin{figure}[h]
  \includegraphics[width=\textwidth]{PredictionsModel1/PredErrBars_Storm46.png}
  \caption{OSA Predictions with $\pm \sigma$ error bars for event: 2003/11/20 to 2003/11/22}
  \label{fig:ComparePred3}
\end{figure}

\section{Conclusion}

In this chapter, we introduced the reader to the \emph{Gaussian
  Process} prediction methodology in the context of space weather forecasting.

In order to successfully apply a Gaussian Process model for a
prediction problem, there are a number of design choices which the
modeler must take. These include choice of model trend, kernel, and
hyperparameter selection among others, we gave a detailed
description of each step of the process from the data to a operational
Gaussian Process forecast model.

To eludicate each step of the modeling workflow, we
considered as an example, one hour ahead prediction of the Dst
index.

We demonstrated that in the one hour ahead Dst prediction
context GP models yield reliable probabilistic predictions with
relatively small data sizes: the training set contains 243
instances, while the validation set contains 782 instances.

To understand how well the models perform, their performance is
compared with the trivial one hour ahead baseline; the
\emph{persistence model}. The GP-AR and GP-ARX both improved on the
performance of the \emph{persistence model}.

\emph{Gaussian Processes} present the data driven space weather practitioner with
a tractable and expressive method for generating probabilistic
forecasts of geomagnetic time series. Leveraging the strengths of the
Bayesian approach, they are able to learn robust predictors from data.

\begin{table}[h]
  \centering
  \caption{Storm events used for model selection of GP-AR and GP-ARX}
  \label{table:validationstorms}
  \begin{tabular}{llllll}
    \hline
    Event Id & Start Date & Start Hour & End Date & End Hour & min. Dst \\ \hline
    1 & 1995/03/26 & 0500 & 1995/03/26 & 2300 & −107 \\
    2 & 1995/04/07 & 1300 & 1995/04/09 & 0900 & −149 \\
    3 & 1995/09/27 & 0100 & 1995/09/28 & 0400 & −108 \\
    4 & 1995/10/18 & 1300 & 1995/10/19 & 1400 & −127 \\
    5 & 1996/10/22 & 2200 & 1996/10/23 & 1100 & −105 \\
    6 & 1997/04/21 & 1000 & 1997/04/22 & 0900 & −107 \\
    7 & 1997/05/15 & 0300 & 1997/05/16 & 0000 & −115 \\
    8 & 1997/10/10 & 1800 & 1997/10/11 & 1900 & −130 \\
    9 & 1997/11/07 & 0000 & 1997/11/07 & 1800 & −110 \\
    10 & 1997/11/22 & 2100 & 1997/11/24 & 0400 & −108 \\
    11 & 2005/06/12 & 1700 & 2005/06/13 & 1900 & −106 \\
    12 & 2005/08/31 & 1200 & 2005/09/01 & 1200 & −122 \\
    13 & 2006/12/14 & 2100 & 2006/12/16 & 0300 & −162 \\
    14 & 2011/09/26 & 1400 & 2011/09/27 & 1200 & −101 \\
    15 & 2011/10/24 & 2000 & 2011/10/25 & 1400 & −132 \\
    16 & 2012/03/08 & 1200 & 2012/03/10 & 1600 & −131 \\
    17 & 2012/04/23 & 1100 & 2012/04/24 & 1300 & −108 \\
    18 & 2012/07/15 & 0100 & 2012/07/16 & 2300 & −127 \\
    19 & 2012/09/30 & 1300 & 2012/10/01 & 1800 & −119 \\
    20 & 2012/10/08 & 0200 & 2012/10/09 & 1700 & −105 \\
    21 & 2012/11/13 & 1800 & 2012/11/14 & 1800 & −108 \\
    22 & 2013/03/17 & 0700 & 2013/03/18 & 1000 & −132 \\
    23 & 2013/05/31 & 1800 & 2013/06/01 & 2000 & −119 \\
    24 & 2014/02/18 & 1500 & 2014/02/19 & 1600 & −112
  \end{tabular}
\end{table}

\begin{table}[h]
  \fontsize{6}{7.5}\selectfont
  \centering
  \caption{Storm events used to evaluate GP-AR and GP-ARX models}
  \label{table:teststorms}
  \begin{tabular}{cccccc}
    \hline
    Event Id & Start Date & Start Time & End Date & End Time & min. Dst \\ \hline
    1 & 1998/02/17 & 1200 & 1998/02/18 & 1000 & -100 \\
    2 & 1998/03/10 & 1100 & 1998/03/11 & 1800 & -116 \\
    3 & 1998/05/04 & 0200 & 1998/05/05 & 0200 & -205 \\
    4 & 1998/08/26 & 1000 & 1998/08/29 & 0700 & -155 \\
    5 & 1998/09/25 & 0100 & 1998/09/26 & 0000 & -207 \\
    6 & 1998/10/19 & 0500 & 1998/10/20 & 0800 & -112 \\
    7 & 1998/11/09 & 0300 & 1998/11/10 & 1600 & -142 \\
    8 & 1998/11/13 & 0000 & 1998/11/15 & 0400 & -131 \\
    9 & 1999/01/13 & 1600 & 1999/01/14 & 2000 & -112 \\
    10 & 1999/02/18 & 0300 & 1999/02/19 & 2100 & -123 \\
    11 & 1999/09/22 & 2000 & 1999/09/23 & 2300 & -173 \\
    12 & 1999/10/22 & 0000 & 1999/10/23 & 1400 & -237 \\
    13 & 2000/02/12 & 0500 & 2000/02/13 & 1500 & -133 \\
    14 & 2000/04/06 & 1700 & 2000/04/08 & 0900 & -288 \\
    15 & 2000/05/24 & 0100 & 2000/05/25 & 2000 & -147 \\
    16 & 2000/08/10 & 2000 & 2000/08/11 & 1800 & -106 \\
    17 & 2000/08/12 & 0200 & 2000/08/13 & 1700 & -235 \\
    18 & 2000/10/13 & 0200 & 2000/10/14 & 2300 & -107 \\
    19 & 2000/10/28 & 2000 & 2000/10/29 & 2000 & -127 \\
    20 & 2000/11/06 & 1300 & 2000/11/07 & 1800 & -159 \\
    21 & 2000/11/28 & 1800 & 2000/11/29 & 2300 & -119 \\
    22 & 2001/03/19 & 1500 & 2001/03/21 & 2300 & -149 \\
    23 & 2001/03/31 & 0400 & 2001/04/01 & 2100 & -387 \\
    24 & 2001/04/11 & 1600 & 2001/04/13 & 0700 & -271 \\
    25 & 2001/04/18 & 0100 & 2001/04/18 & 1300 & -114 \\
    26 & 2001/04/22 & 0200 & 2001/04/23 & 1500 & -102 \\
    27 & 2001/08/17 & 1600 & 2001/08/18 & 1600 & -105 \\
    28 & 2001/09/30 & 2300 & 2001/10/02 & 0000 & -148 \\
    29 & 2001/10/21 & 1700 & 2001/10/24 & 1100 & -187 \\
    30 & 2001/10/28 & 0300 & 2001/10/29 & 2200 & -157 \\
    31 & 2002/03/23 & 1400 & 2002/03/25 & 0500 & -100 \\
    32 & 2002/04/17 & 1100 & 2002/04/19 & 0200 & -127 \\
    33 & 2002/04/19 & 0900 & 2002/04/21 & 0600 & -149 \\
    34 & 2002/05/11 & 1000 & 2002/05/12 & 1600 & -110 \\
    35 & 2002/05/23 & 1200 & 2002/05/24 & 2300 & -109 \\
    36 & 2002/08/01 & 2300 & 2002/08/02 & 0900 & -102 \\
    37 & 2002/09/04 & 0100 & 2002/09/05 & 0000 & -109 \\
    38 & 2002/09/07 & 1400 & 2002/09/08 & 2000 & -181 \\
    39 & 2002/10/01 & 0600 & 2002/10/03 & 0800 & -176 \\
    40 & 2002/10/03 & 1000 & 2002/10/04 & 1800 & -146 \\
    41 & 2002/11/20 & 1600 & 2002/11/22 & 0600 & -128 \\
    42 & 2003/05/29 & 2000 & 2003/05/30 & 1000 & -144 \\
    43 & 2003/06/17 & 1900 & 2003/06/19 & 0300 & -141 \\
    44 & 2003/07/11 & 1500 & 2003/07/12 & 1600 & -105 \\
    45 & 2003/08/17 & 1800 & 2003/08/19 & 1100 & -148 \\
    46 & 2003/11/20 & 1200 & 2003/11/22 & 0000 & -422 \\
    47 & 2004/01/22 & 0300 & 2004/01/24 & 0000 & -149 \\
    48 & 2004/02/11 & 1000 & 2004/02/12 & 0000 & -105 \\
    49 & 2004/04/03 & 1400 & 2004/04/04 & 0800 & -112 \\
    50 & 2004/07/22 & 2000 & 2004/07/23 & 2000 & -101 \\
    51 & 2004/07/24 & 2100 & 2004/07/26 & 1700 & -148 \\
    52 & 2004/07/26 & 2200 & 2004/07/30 & 0500 & -197 \\
    53 & 2004/08/30 & 0500 & 2004/08/31 & 2100 & -126 \\
    54 & 2004/11/07 & 2100 & 2004/11/08 & 2100 & -373 \\
    55 & 2004/11/09 & 1100 & 2004/11/11 & 0900 & -289 \\
    56 & 2004/11/11 & 2200 & 2004/11/13 & 1300 & -109 \\
    57 & 2005/01/21 & 1800 & 2005/01/23 & 0500 & -105 \\
    58 & 2005/05/07 & 2000 & 2005/05/09 & 1000 & -127 \\
    59 & 2005/05/29 & 2200 & 2005/05/31 & 0800 & -138 \\
    60 & 2005/06/12 & 1700 & 2005/06/13 & 1900 & -106 \\
    61 & 2005/08/31 & 1200 & 2005/09/01 & 1200 & -131 \\
    62 & 2006/04/13 & 2000 & 2006/04/14 & 2300 & -111 \\
    63 & 2006/12/14 & 2100 & 2006/12/16 & 0300 & -147 \\ \hline
  \end{tabular}%
\end{table}

\begin{table}[h]
  \centering
  \caption{Settings of model selection procedures}
  \begin{tabular}{l c c c}
    \hline
    Procedure & Grid Size & Step & Max Iterations \\
    \hline
    Grid Search & 10 & 0.2 & NA \\
    Coupled Simulated Annealing & 4 & 0.2 & 30 \\
    Maximum likelihood & NA & 0.2 & 150\\
  \end{tabular}
  \label{table:modelselection}
\end{table}

\begin{table}[h]
  \centering
  \caption{Evaluation results for models on storm events listed in table \ref{table:teststorms}}
  \label{table:results}
  \begin{tabular}{l c c c}
    \hline
    Model & Mean Absolute Error & Root Mean Square Error & Coefficient of Correlation\\ \hline
    GP-ARX & 7.252 & 11.93 & 0.972\\
    GP-AR & 8.37 & 14.04 & 0.963\\
    Persistence & 9.182 & 14.94 & 0.957\\
  \end{tabular}
\end{table}

\clearpage

\bibliography{references}

\end{document}